---
title: "Activity Data"
author: "Me"
date: "7/07/2021"
output:
  html_document: default
  pdf_document: default
---


Loads data and replaces NA values with null value.


```{r activity_impute_nas, echo=TRUE}
activity <- read.csv("activity.csv", header=TRUE)
activity$interval <- as.numeric(activity$interval)
activity$steps <- as.numeric(activity$steps)
activity$date <- as.Date(activity$date)
activity[is.na(activity)] <- 0
```


Groups and sums steps taken over 5 minute intervals by date, plots histogram of daily steps, and takes summary statistics to calculate mean and median of daily steps taken. 


```{r activity_steps, echo=TRUE}
library(dplyr)
activity_steps <- 
  activity %>% 
  group_by(date) %>% 
  summarise(steps = sum(steps))
activity_steps$date <- as.Date(activity_steps$date)
hist(activity_steps$steps, 
    main="Histogram of Daily Steps", 
    xlab="Number of Daily Steps", ylim=c(0,30))
summary(activity_steps$steps)
```


Calculates average number of steps for each interval, plots time series of average steps per interval, and calculates 5 minute interval with maximum average number of daily steps.


```{r plot avg_steps_interval, echo=TRUE}
avg_steps_int <- group_by(activity, interval) %>% summarize(m = mean(steps))
names(avg_steps_int)[2] <- "steps"
plot(avg_steps_int$interval, 
    avg_steps_int$steps, type = "l", 
    main="Average Steps Per Interval", col=2, 
    xlab="5 minute interval", ylab="average number of steps", 
    ylim=c(0,200))
max_int <- which.max(avg_steps_int$interval) *5
max_int
```


Replaces NA values of "steps" with mean of "steps" column, plots histogram of daily steps, and takes summary statistics to calculate mean and median of daily steps taken.


```{r impute NAs, echo=TRUE}
activity <- read.csv("activity.csv", header=TRUE)
sum(is.na(activity$steps))

library(dplyr)
activity_impute <- activity %>%
  group_by(interval) %>%
  mutate(steps=ifelse(is.na(steps), mean(steps,na.rm=TRUE), steps))
activity_impute_2 <- activity_impute %>%
  group_by(date) %>%
  summarise(steps=sum(steps))
hist(activity_impute_2$steps, 
     main="Histogram of Daily Steps", 
     xlab="Number of Daily Steps", 
     ylim=c(0,40))
summary(activity_impute_2$steps)
```


Categorizes dates as weekdays and weekend days. Averages steps according to these categories. Creates panel plots to compare average step patterns across 5 minute intervals for weekdays and weekend days.


```{r separate weekdays and weekends, echo=TRUE}
activity <- read.csv("activity.csv", header=TRUE)
activity_impute <- activity %>%
  group_by(interval) %>%
  mutate(steps=ifelse(is.na(steps), mean(steps,na.rm=TRUE), steps))

activity_impute$interval <- as.numeric(activity_impute$interval)
activity_impute$steps <- as.numeric(activity_impute$steps)
activity_impute$date <- as.Date(activity_impute$date)

library(dplyr)
activity_impute$day <- weekdays(activity_impute$date)
weekend <- c("Saturday", "Sunday")
activity_impute$category[activity_impute$day==weekend] <- "weekend"
activity_impute$category[activity_impute$day!=weekend] <- "weekday"

avg_steps_day <- group_by(activity_impute, interval, category) %>% 
  summarize(m = mean(steps))
names(avg_steps_day)[3] <- "steps"

library(ggplot2)
p <- ggplot(data = avg_steps_day, aes(x = interval, y = steps)) + geom_line()
p + facet_wrap(~category)

```